---
title: "EDA_2019_01_29"
author: "Andrew Srisuwananukorn"
date: "2/2/2019"
output: html_document
---

#Introduction
- exploration of cheese data provided by Thomas Mock of Tidy Tuesday
https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-01-29

- Also inspired by clustering methods DataCamp lesson by Dmitry 

#Import libraries
```{r}
library(tidyverse)
library(ggplot2)
library(purrr)
library(cluster)
library(dendextend)

```


#Import Data
```{r}
raw_cheese_df <- read_csv("/Users/Andrewsris/Box/Tidy_Tuesday/2019_01_29/data/1_original/clean_cheese.csv") 
cheese_df <- raw_cheese_df %>% 
  gather(key = Cheese, value = value, 2:ncol(raw_cheese_df)) %>% 
  spread(key = names(raw_cheese_df)[1], value = "value") %>% 
  filter(!str_detect(Cheese, "Total"))
```

#Scale data
This is unneceesary for this specific data frame, as the scale is the same for each feature. However I am including this section for future datasets, when i use similar methods.

```{r}
scaled_cheese_df <- cheese_df %>% column_to_rownames("Cheese") %>% scale() 
scaled_dist_cheese <- scaled_cheese_df %>% dist()
```

#Hierarchical clustering
Distance = "euclidian"
hierarchical clustering method = "complete"
```{r}
hc_cheese <- hclust(scaled_dist_cheese, method = "complete")

#By visual inspection, optimal k=2
hc_cheese %>% as.dendrogram() %>% set("branches_k_color", h = 10) %>% plot()
par(mar = c(5,4,1,6)) #Reset margins: bottom, left, top, right
hc_cheese %>% as.dendrogram() %>% 
  set("branches_k_color", k = 2) %>% 
  set("labels_cex", 0.75) %>% 
  plot(horiz = TRUE,
       main = "Hierarchical Clustering by Cheese Consumption", 
       sub = "k=2 by visual inspection") %>% 
  abline(v = 15, lty = 2)

#Assign cluster groups
clusters_hc <- cutree(hc_cheese, k = 2)

```

#K-means analysis
```{r}
#We will use the elbow plot to determine optimal k
# Use map_dbl to run many models with varying value of k (centers)
tot_withinss <- map_dbl(1:10,  function(k){
  model <- kmeans(x = scaled_dist_cheese, centers = k)
  model$tot.withinss
})

# Generate a data frame containing both k and tot_withinss
elbow_df <- data.frame(
  k = 1:10 ,
  tot_withinss = tot_withinss
)
ggplot(elbow_df, aes(x = k , y = tot_withinss)) +
  geom_line() +
  scale_x_continuous(breaks = 1:10) + 
  labs(title = "Elbow Plot for K-means clustering of Cheese Consumption",
       subtitle = "Optimal k=2 by visual inspection of elbow point",
       y = "Total within squared sums",
       x = "k") +
  geom_point(data = elbow_df %>% filter(k == 2), aes(x = k, y = tot_withinss, color = "red", size = 3), show.legend = FALSE)

#Optimal k = 2
km_cheese <- kmeans(scaled_dist_cheese, centers = 2)
clusters_km <- km_cheese$cluster


```

#PAM and silhouette
- pam is from the cluster pacakage
```{r}
#using the silhouette technique to determine optimal k

# Use map_dbl to run many models with varying value of k
sil_width <- map_dbl(2:10,  function(k){
  model <- pam(x = scaled_dist_cheese, k = k)
  model$silinfo$avg.width
})

# Generate a data frame containing both k and sil_width
sil_df <- data.frame(
  k = 2:10,
  sil_width = sil_width
)

# Plot the relationship between k and sil_width
ggplot(sil_df, aes(x = k, y = sil_width)) +
  geom_line() +
  scale_x_continuous(breaks = 2:10) +   
  labs(title = "Silhouette Plot for K-means clustering of Cheese Consumption",
       subtitle = "Optimal k=2 by maximal silhouette width",
       y = "Average silhouette width",
       x = "k") +
  geom_point(data = sil_df %>% filter(k == 2), aes(x = k, y = sil_width, color = "red", size = 3), show.legend = FALSE)

#Optimal k = 2
pam_cheese <- pam(scaled_dist_cheese, k = 2)
plot(silhouette(pam_cheese))

clusters_pam <- pam_cheese$clustering

```


#Merge df
```{r}
total_cheese <- mutate(cheese_df, 
               cluster_hc = clusters_hc,
               cluster_km = clusters_km,
               cluster_pam = clusters_pam) %>% gather(-c("Cheese","cluster_km", "cluster_hc", "cluster_pam"), key = "year", value = "eaten")

total_cheese %>% ggplot(aes(x = year, y = eaten, color = factor(cluster_km))) + 
  geom_line(aes(group = Cheese),show.legend = FALSE) + 
  geom_smooth(aes(group = cluster_km), na.rm=TRUE, show.legend = FALSE) +
  geom_text(data = total_cheese %>% filter(year == 2013), 
            aes(label = Cheese), size = 2.5, nudge_y = 0.5, show.legend = FALSE) +
  scale_x_discrete(breaks = seq(1970,2015, 5)) +
  theme(axis.text.x = element_text(angle = 45)) +
  labs(title = "Average cheese consumption from USDA",
       subtitle = "Optimal k=2 by Elbow method of K-means cluster analysis",
       y = "Average cheese consumption (lbs per person)",
       x = "Year"
       )
```

