---
title: "Medium Data Science Articles"
author: "Andrew Srisuwananukorn"
date: "12/4/2018"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
```

## Upload the data

```{r}
df <- read_csv("~/Box/Tidy_Tuesday/2018_12_4/data/1_original/medium_datasci.csv")

```

#Brief exploration
```{r}
glimpse(df)

skimr::skim(df)
```

#Explore Tag columns
Are the tags mutually exclusive?

```{r}
tags_df <- df %>% select(starts_with("tag"))
num_tags <- tags_df %>% 
  summarise_at(1:ncol(.), ~sum(.)) %>% 
  gather(tag, n) %>% 
  mutate(tag = str_remove(tag, "tag_")) %>% 
  arrange(desc(n))
num_tags
```

#Gather Tags example
```{r}
tags2_df <- df %>% select("x1", starts_with("tag"))
tags2_df
tags2_df %>% gather(tag, value, -x1)

# tags2_df %>% select(-starts_with("tag"))
tags2_df %>% gather(tag, value, starts_with("tag"))
```


#Tidying data
Here are the column names
```{r}
colnames(df)
```


```{r}
tidy_df <- df  %>% 
  unite("date", c("month", "day", "year"), remove = TRUE) %>% 
  mutate(date = mdy(date)) %>% 
  gather(tag, tag_value, starts_with("tag")) %>% 
  mutate(tag = str_remove(tag, "tag_")) %>% 
  filter(tag_value == 1) %>% 
  select(-tag_value)

tidy_df
```

#Top ten Author Plots
```{r}
tidy_df %>% 
  count(author) %>% 
  filter(!is.na(author)) %>% 
  arrange(desc(n)) %>% head(n=10) %>% 
  ggplot(aes(x = fct_reorder(author,n), y = n)) + 
  geom_point() + 
  coord_flip() + 
  ggtitle("Top Ten Authors") + 
  labs(x = "Author", y = "# of Articles" )
```

## Plot histogram based on date
Interesting to note that there are much fewer articles in December
```{r}
tidy_df %>% ggplot(aes(x=date, color=tag)) + geom_histogram(binwidth=7) + scale_x_date(date_breaks="month") + theme(axis.text.x=element_text(angle=45, hjust=1)) + ggtitle("Articles by Tag")

tidy_df %>% ggplot(aes(x=date, color=tag)) + geom_freqpoly(binwidth=7) + scale_x_date(date_breaks = "month") + 
theme(axis.text.x=element_text(angle=45, hjust=1)) + ggtitle("Articles by Tag")
```

#Focus on December
```{r}
tidy_df %>% filter(date >= mdy("11_1_17"), date <= mdy("1_31_18")) %>% ggplot(aes(x=date, color=tag)) + geom_bar(binwidth=1) + scale_x_date(date_breaks = "week") + theme(axis.text.x=element_text(angle=45, hjust=1)) + ggtitle("Articles In December 2017")
```

#Addendum
After posting, Nathanmoore @nmoorenz suggested that I use fill=tag instead of colour=tag

```{r}
tidy_df %>% ggplot(aes(x=date, fill=tag)) + geom_histogram(binwidth=7) + scale_x_date(date_breaks="month") + theme(axis.text.x=element_text(angle=45, hjust=1)) + ggtitle("Articles by Tag")

tidy_df %>% filter(date >= mdy("11_1_17"), date <= mdy("1_31_18")) %>% ggplot(aes(x=date, fill=tag)) + geom_bar(binwidth=1) + scale_x_date(date_breaks = "week") + theme(axis.text.x=element_text(angle=45, hjust=1)) + ggtitle("Articles In December 2017")
```


